---
title: "Development and Internal Validation of a Model to Predict Malignancy of Adnexal Masses"
subtitle: "Part 1: Development and Validation"
author: 
  - "Group D"
date: today
execute: 
  echo: true
  warning: false
format:
  docx:
    reference-doc: ../docs/manuscript/template.docx
    link-citations: true
  pdf:
    toc: true
editor: visual
---

{{< pagebreak >}}

```{r}
#| label: directories
#| include: false

# Run only when using Positron (for compatibility with RStudio project)
# setwd(paste0(getwd(),"/R"))

# Create directories for sub-folders  
inputfolder <- "../data/raw"
psfolder <- "../data/processed"
tempfolder <- "../data/temp"
figfolder <- "../results/output_figures"
tabfolder <- "../results/output_tables"

dir.create(inputfolder, showWarnings = FALSE)
dir.create(psfolder, showWarnings = FALSE)
dir.create(tempfolder, showWarnings = FALSE)
dir.create(figfolder, showWarnings = FALSE)
dir.create(tabfolder, showWarnings = FALSE)
```

```{r}
#| label: load packages
#| include: false

if (!require("pacman", quietly = TRUE)) {
  install.packages("pacman")
}

pacman::p_load(
  tidyverse,  
  pROC,
  rms,
  rmda,
  gt,         
  report     
)

```

```{r}
#| label: load data
#| include: false

data <- read.delim(
  file.path(inputfolder,"ovarian.txt")
  )

# Add labels for variables in the model
Hmisc::label(data$Age) <- "Age (years)"
Hmisc::label(data$Ascites) <- "Ascites"
Hmisc::label(data$papflow) <- "Papillations with blood flow"
Hmisc::label(data$soldmaxorig) <- "Maximum diameter of solid component (mm)"
Hmisc::label(data$wallreg) <- "Internal cyst wall regularity"
Hmisc::label(data$Shadows) <- "Acoustic shadows"
Hmisc::label(data$mal) <- "Tumor classification"
Hmisc::label(data$CA125) <- "CA-125"
```

```{r}
#| include: false
options(
  datadist=NULL,
  prType='html'
  )

dd <- datadist(data); options(datadist='dd')
```

::: landscape

# Model Development

```{r}
#| label: dev fit model 1

model_1 <- lrm(
  mal ~ rcs(Age,3) + Ascites + papflow + rcs(soldmaxorig,3) + wallreg + Shadows,
  data = data,
  x=TRUE, y=TRUE
)
```

```{r}
#| echo: false
summary(model_1, 
  vnames = "labels"
)
```

::: 

```{r}
#| label: fig-nonlinear
#| fig-cap: "Visualization of non-linear terms"
#| fig-subcap: 
#|   - "Age (k = 3)"
#|   - "Maximum diameter of solid component (k = 3)"
#| layout-nrow: 2
#| fig-pos: "H"
#| echo: false

plot(Predict(model_1, Age))
plot(Predict(model_1, soldmaxorig))
```

{{< pagebreak >}}

```{r}
anova(model_1)
```

A non-linear term for age is not better than a linear term. Thus, we will model age as a linear predictor. For the maximum diameter of solid component (mm), a non-linear term is clearly better than a linear term. 


```{r}
#| label: dev fit model 2

model_2 <- lrm(
  mal ~ Age + Ascites + papflow + rcs(soldmaxorig,3) + wallreg + Shadows,
  data = data,
  x=TRUE, y=TRUE
)
```

::: landscape

```{r}
#| echo: false
summary(model_2, 
  vnames = "labels"
)
```

```{r}
#| label: extract-coefs
#| include: false

b <- coef(model_2)

# Verify exact names with: names(b)
b0 <- b["Intercept"]
b1 <- b["Age"]
b2 <- b["Ascites"]    
b3 <- b["papflow"]
b4 <- b["soldmaxorig"]  
b5 <- b["soldmaxorig'"]  
b6 <- b["wallreg"]
b7 <- b["Shadows"]

```

The log odds of malignancy was modeled as:

$$
\begin{aligned}
\text{log odds} = \; & \beta_0 + \beta_1 \cdot \text{Age} + \beta_2 \cdot \text{Ascites} 
+ \beta_3 \cdot \text{Papillations with flow} + \beta_4 \cdot \text{Solid diameter} \\
& + \beta_5 \cdot \text{Solid diameter}' + \beta_6 \cdot \text{Irregular wall}
+ \beta_7 \cdot \text{Acoustic shadows}
\end{aligned}
$$

where $\text{Solid diameter}'$ is the restricted cubic spline transformation term. Substituting the estimated coefficients, the linear predictor (LP) is:

$$
\begin{aligned}
\text{LP} = \; & `r round(b0, 3)`
+ `r round(b1, 3)` \cdot \text{Age}
+ `r round(b2, 3)` \cdot \text{Ascites} 
+ `r round(b3, 3)` \cdot \text{Papillations with flow}
+ `r round(b4, 3)` \cdot \text{Solid diameter} \\
& + `r round(b5, 3)` \cdot \text{Solid diameter}'
+ `r round(b6, 3)` \cdot \text{Irregular wall}
+ `r round(b7, 3)` \cdot \text{Acoustic shadows}
\end{aligned}
$$

::: 

# Model Validation

## Internal Validation 

```{r}
# Bootstrap validation
validate_results <- validate(model_2, method = "boot", B = 200)
validate_results
```

## Discrimination

```{r}
# Apparent C-statistic
c_apparent  <- 0.5 * (validate_results["Dxy", "index.orig"] + 1)
c_apparent

# Optimism-corrected C-statistic
c_corrected <- 0.5 * (validate_results["Dxy", "index.corrected"] + 1)
c_corrected
```

```{r}
#ROC for the model
#get predicted probabilities from development model
pred_probs <- predict(model_2, type="fitted")

roc_apparent <- roc(data$mal, pred_probs)

roc_dataframe <- data.frame(
  specificity = roc_apparent$specificities, 
  sensitivity = roc_apparent$sensitivities
  )

#plot
auc_val <- auc(roc_apparent)
auc_ci  <- ci.auc(roc_apparent, conf.level = 0.95)
auc_label <- paste0(
  "AUC = ", round(auc_val, 2), 
  "\n95% CI: ", round(auc_ci[1], 2), 
  " \u2013 ", round(auc_ci[3], 2)
  )

ggplot(roc_dataframe, aes(x = 1 - specificity, y = sensitivity)) + 
  geom_line(color = "black") +
  geom_abline(linetype = "dashed", color = "gray") + 
  scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.2)) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.2)) + 
  annotate("text", x = 0.75, y = 0.25, label = auc_label, size = 5) +
  labs(x = "1 - Specificity", y = "Sensitivity") + theme_classic(base_size = 14)
```

{{< pagebreak >}}

```{r}
#diagnostic accuracy for 5%, 10%, 15% and 20% risk cut-offs
cutoffs <- c(0.05, 0.10, 0.15, 0.20)

#function to calculate diagnostic accuracy cut-offs
diag_accuracy <- function(probs, outcome, cutoff) {
  pred <- ifelse(probs >= cutoff, 1, 0)
  
  TP <- sum(pred == 1 & outcome == 1)
  TN <- sum(pred == 0 & outcome == 0)
  FP <- sum(pred == 1 & outcome == 0)
  FN <- sum(pred == 0 & outcome == 1)
  
  sens <- TP / (TP + FN)
  spec <- TN / (TN + FP)
  ppv  <- TP / (TP + FP)
  npv  <- TN / (TN + FN)
  
  ci_sens <- prop.test(TP, TP + FN, correct = FALSE)$conf.int
  ci_spec <- prop.test(TN, TN + FP, correct = FALSE)$conf.int
  ci_ppv  <- prop.test(TP, TP + FP, correct = FALSE)$conf.int
  ci_npv  <- prop.test(TN, TN + FN, correct = FALSE)$conf.int
  
  fmt <- function(est, ci) {
    sprintf("%.3f (%.3f\u2013%.3f)", est, ci[1], ci[2])
  }
  
  data.frame(
    Cutoff      = cutoff,
    Sensitivity = fmt(sens, ci_sens),
    Specificity = fmt(spec, ci_spec),
    PPV         = fmt(ppv,  ci_ppv),
    NPV         = fmt(npv,  ci_npv)
  )
}

results_diag_accuracy <- do.call(
  rbind, 
  lapply(cutoffs, diag_accuracy, probs = pred_probs, outcome = data$mal)
  )
```

::: landscape

```{r}
#| echo: false  
results_diag_accuracy %>% gt
```

::: 

## Calibration 

```{r}
# Calibration slope
round(validate_results["Slope", "index.corrected"], 4)

# Calibration plot
calibration_results <- calibrate(model_2, B = 200)
```

```{r}
plot(
  calibration_results, 
  lwd = 2, xlim = c(0, 1), ylim = c(0, 1), 
  xlab = "Predicted Probability", ylab = "Observed Probability", 
  subtitles = FALSE, legend=FALSE
  )
legend(
  "bottomright", 
  legend = c("Ideal", "Apparent", "Optimism-corrected", "95% CI"), 
  col = c("gray60", "gray40", "black", "gray85"), 
  lty = c(2, 1, 1, 1), lwd = c(2, 2, 2, 1), bty = "n"
  )
```

{{< pagebreak >}}

# Clinical Utility

```{r}
data <- cbind(data, pred_probs) 

dca <- decision_curve(
  mal ~ pred_probs,
  data = data,
  fitted.risk=TRUE
  )

plot_decision_curve(
  dca, 
  curve.names = c("Malignancy model", "Treat all", "Treat none")
  )
```


{{< pagebreak >}}

# Package References

```{r}
#| echo: false  

# remove clutter
session <- sessionInfo()
session$BLAS <- NULL
session$LAPACK <- NULL
session$loadedOnly <- NULL
```

```{r}
#| output: asis
#| echo: false  
report::cite_packages(session)
```

```{r}
#| include: false

# Run this chunk if you wish to clear your environment and unload packages.

pacman::p_unload(negate = TRUE)

rm(list = ls())
```
